### 2025-07-24 draft-2/draft-2.tex
began work on a few items

\includeruby color.rb
\includeruby syntax.rb

@@

\beginruby

@@

\endruby

@@

  # Pseudocode
  amber_fg = lch(0.7, 0.125,  75)
  green_fg = lch(0.6, 0.15 , 150)
   blue_fg = lch(0.5, 0.2  , 270)
  lilac_fg = lch(0.6, 0.2  , 300)
\endruby
%
\newtoks\restorecolor \restorecolor={0 0 0}%
%
\long\def\color#1#2{%
  {\restorecolor={#1}\special{color rgb #1}#2}%
  \special{color rgb \the\restorecolor}%
}%
%
\long\def\colorbox#1#2#3{%
  \par
  % Set the contents of the box
  \advance\hsize by -2em
  \setbox0 = \vbox{#3}%
  \advance\hsize by 2em
  % Insert the usual interline glue
  \ifdim \prevdepth > -1000pt
    \vskip-\prevdepth
    \vskip\baselineskip
  \fi
  % Insert an extra half of a baselineskip above
  \vskip0.5\baselineskip
  % Calculate dimensions
  \dimen0 = \ht0
  \dimen1 = \dp0
  \advance\dimen0 by 0.25\baselineskip % midway along half-baselineskip above
  \advance\dimen0 by 0.25\baselineskip % adjust the upper edge and
  \advance\dimen1 by 0.20\baselineskip %   lower edge to be perceptually correct
  \advance\dimen1 by 0.25\baselineskip % midway along half-baselineskip below
  % Insert the background
  \kern -0.25\baselineskip % midway along extra
  \kern -0.25\baselineskip % adjustment
  \color{#1}{\hrule width \hsize height \dimen0 depth \dimen1}%
  \kern -\ht0
  \kern -\dp0
  \kern -0.20\baselineskip % adjustment
  \kern -0.25\baselineskip % midway along extra
  % Insert the ribbon
  \kern -0.25\baselineskip % midway along extra
  \kern -0.25\baselineskip % adjustment
  \color{#2}{\hrule width 0.5em height \dimen0 depth \dimen1}%
  \kern -\ht0
  \kern -\dp0
  \kern -0.20\baselineskip % adjustment
  \kern -0.25\baselineskip % midway along extra
  % Insert the box
  \hbox{\hskip1em\box0}
  % Insert an extra half of a baselineskip below
  \vskip0.5\baselineskip
}%
\long\def  \greybox#1{\colorbox{\r{  grey_bg}}{\r{  grey_rbn}}{#1}}%
\long\def\purplebox#1{\colorbox{\r{purple_bg}}{\r{purple_rbn}}{#1}}%
\long\def  \bluebox#1{\colorbox{\r{  blue_bg}}{\r{  blue_rbn}}{#1}}%
\long\def \greenbox#1{\colorbox{\r{ green_bg}}{\r{ green_rbn}}{#1}}%
\long\def\yellowbox#1{\colorbox{\r{yellow_bg}}{\r{yellow_rbn}}{#1}}%
\long\def   \redbox#1{\colorbox{\r{   red_bg}}{\r{   red_rbn}}{#1}}%
%
\long\def\label  #1{\color{\r{ label_fg}}{#1}}%
\long\def\term   #1{\color{\r{purple_fg}}{#1}}%
\long\def\comment#1{\color{\r{  grey_fg}}{{\noto ※ }#1}}%
\long\def\s      #1{\color{\r{  teal_fg}}{\fw #1}}%
%
\def\hex#1{{\fw\color{\r{ltgrey_fg}}{\cas x}#1}}%
\def\uni#1{{\caps u+#1}}%
%
\def\ellipsis{.\ot.\ot.}
%
\newcount\majorcount
\newcount\minorcount
\def\major{%
  \advance\majorcount by 1
  \minorcount=0
  \leavevmode
  \llap{\label{\the\majorcount}\hem}%
}%
\def\minor{%
  \advance\minorcount by 1
  \leavevmode
  \llap{\label{\the\majorcount·\the\minorcount}\hem}%
}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\gets{<-}%
\def\to  {->}%
\def\then{=>}%
%
\setbox0=\hbox{\fw ()[]\char"7B\char"7D\it ()[]\char"7B\char"7D}%
\newdimen\strutht
\newdimen\strutdp
\strutht=\ht0
\strutdp=\dp0
\dimen0=\baselineskip
\advance\dimen0 by -\strutht
\advance\dimen0 by -\strutdp
\divide\dimen0 by 2
\advance\strutht by \dimen0
\advance\strutdp by \dimen0
\def\strut{\vrule width 0pt height \strutht depth \strutdp}%
%
\def\tab{%
  \kern 0.75ex
  \color{\r{ltgrey_fg}}{\vrule width 0.5pt}%
  \kern -0.5pt
  \kern -0.75ex
  \kern 3ex
}%
%
\def\kwd  #1{\color{\r{  blue_fg}}{#1}}% def, return, break, error
\def\flow #1{\color{\r{ amber_fg}}{#1}}% repeat, if, then, match
\def\liter#1{\color{\r{   red_fg}}{#1}}% numeric
\def\const#1{\color{\r{ lilac_fg}}{#1}}% empty, eof
\def\func #1{\color{\r{ green_fg}}{#1}}%
\def\var  #1{{\ifw #1}}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\centerline{\lcap8 UNIVERSAL CHESS INTERFACE}
\centerline{\lcap4 november 2024 • draft}

@@

\vskip-0.5\parskip The final position (\ot{\it P}\isub{N} if \s{moves} is
specified, otherwise {\it P}\sub{0}\ot) is the position \term{described} by
the \s{position} message.

% TODO state machine: position message “sets” “current position” to P_N

@@

@@

@@

\minor In some states, the engine may dequeue a message. If the engine dequeues
a message, the message must be well-formed and have an allowed message type
given the state of the engine.
\vskip-\parskip
\greybox{This is not meant to imply that standard input is ever required to be
empty, but meant to indicate that in some states the engine is not allowed to
read standard input (or at least must behave as if it has not read standard
input).}

@@

\minor The client must enqueue a \s{position} message at least once before it
enqueues a \s{go} message for the first time after enqueuing \s{uci}.
\vskip-\parskip
\bluebox{Engines are strongly recommended, if a \s{position} message is not
sent before the first \s{go} message, to behave as if \s{position startpos}
were sent.}
\vskip-\parskip
\greybox{Note that the client is not required to send \s{isready} before
\s{go}.}
\vskip-\parskip
\greybox{Note that the client may send \s{ucinewgame} after \s{position}
before \s{go} without sending any additional intervening \s{position} message.}
\vskip-\parskip
\greybox{Note that \s{ucinewgame} does not reset the current board state nor
does it reset the move history. (Only \s{position} changes the current board
state and move history.)}

@@

\vskip\parskip
\beginruby
format("let @read_byte! : File_Descriptor -> Byte | EOF")
\endruby

\vskip\parskip
\beginruby
format(%q{
def withoutCR('msg : List(Byte)) : List(Byte)
  if empty?('msg) or last('msg) ≠ `0d then return 'msg
  return withoutLast('msg)
})
\endruby

\vskip\parskip
\beginruby
format(%q{
def read_message!(mut 'fd : File_Descriptor) : List(Byte) | EOF
  'msg <- empty
  repeat
    match read_byte!('fd)
      eof => return (\pt if empty?('msg) then eof else 'msg\pt)
      `0a => return withoutCR('msg)
      'val => append!('msg, 'val\pt)
})
\endruby

\vskip\parskip
\beginruby
format(%q{
def get_token!(mut 'msg : List(Byte)) : List(Byte) | None
  repeat
    if empty?('msg) then return none
    if first('msg) ≠ `20 then break
    removeFirst!('msg)
  'tok <- empty
  repeat
    append!('tok, first('msg))
    removeFirst!('msg)
    if empty?('msg) then return 'tok
    if first('msg) = `20 then break
  return 'tok
})
\endruby

>>>>

\include color.rb
\include syntax.rb

@@

\begin

@@

\end

@@

  # Pseudocode
  amber_fg = lch(0.7, 0.125,  75)
  green_fg = lch(0.6, 0.15 , 150)
   blue_fg = lch(0.5, 0.2  , 270)
  lilac_fg = lch(0.6, 0.2  , 300)
\end
%
\newtoks\restorecolor \restorecolor={0 0 0}%
%
\long\def\color#1#2{%
  {\restorecolor={#1}\special{color rgb #1}#2}%
  \special{color rgb \the\restorecolor}%
}%
%
\long\def\colorbox#1#2#3{%
  \par
  % Set the contents of the box
  \advance\hsize by -2em
  \setbox0 = \vbox{#3}%
  \advance\hsize by 2em
  % Insert the usual interline glue
  \ifdim \prevdepth > -1000pt
    \vskip-\prevdepth
    \vskip\baselineskip
  \fi
  % Insert an extra half of a baselineskip above
  \vskip0.5\baselineskip
  % Calculate dimensions
  \dimen0 = \ht0
  \dimen1 = \dp0
  \advance\dimen0 by 0.25\baselineskip % midway along half-baselineskip above
  \advance\dimen0 by 0.25\baselineskip % adjust the upper edge and
  \advance\dimen1 by 0.20\baselineskip %   lower edge to be perceptually correct
  \advance\dimen1 by 0.25\baselineskip % midway along half-baselineskip below
  % Insert the background
  \kern -0.25\baselineskip % midway along extra
  \kern -0.25\baselineskip % adjustment
  \color{#1}{\hrule width \hsize height \dimen0 depth \dimen1}%
  \kern -\ht0
  \kern -\dp0
  \kern -0.20\baselineskip % adjustment
  \kern -0.25\baselineskip % midway along extra
  % Insert the ribbon
  \kern -0.25\baselineskip % midway along extra
  \kern -0.25\baselineskip % adjustment
  \color{#2}{\hrule width 0.5em height \dimen0 depth \dimen1}%
  \kern -\ht0
  \kern -\dp0
  \kern -0.20\baselineskip % adjustment
  \kern -0.25\baselineskip % midway along extra
  % Insert the box
  \hbox{\hskip1em\box0}
  % Insert an extra half of a baselineskip below
  \vskip0.5\baselineskip
}%
\long\def  \greybox#1{\colorbox{\r{  grey_bg}}{\r{  grey_rbn}}{#1}}%
\long\def\purplebox#1{\colorbox{\r{purple_bg}}{\r{purple_rbn}}{#1}}%
\long\def  \bluebox#1{\colorbox{\r{  blue_bg}}{\r{  blue_rbn}}{#1}}%
\long\def \greenbox#1{\colorbox{\r{ green_bg}}{\r{ green_rbn}}{#1}}%
\long\def\yellowbox#1{\colorbox{\r{yellow_bg}}{\r{yellow_rbn}}{#1}}%
\long\def   \redbox#1{\colorbox{\r{   red_bg}}{\r{   red_rbn}}{#1}}%
%
\long\def\label  #1{\color{\r{ label_fg}}{#1}}%
\long\def\term   #1{\color{\r{purple_fg}}{#1}}%
\long\def\comment#1{\color{\r{  grey_fg}}{{\noto ※ }#1}}%
\long\def\s      #1{\color{\r{  teal_fg}}{\fw #1}}%
%
\def\hex#1{{\fw\color{\r{ltgrey_fg}}{\cas x}#1}}%
\def\uni#1{{\caps u+#1}}%
%
\def\ellipsis{.\ot.\ot.}
%
\newcount\majorcount
\newcount\minorcount
\def\major{%
  \advance\majorcount by 1
  \minorcount=0
  \leavevmode
  \llap{\label{\the\majorcount}\hem}%
}%
\def\minor{%
  \advance\minorcount by 1
  \leavevmode
  \llap{\label{\the\majorcount·\the\minorcount}\hem}%
}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\gets{<-}%
\def\to  {->}%
\def\then{=>}%
%
\setbox0=\hbox{\fw ()[]\char"7B\char"7D\it ()[]\char"7B\char"7D}%
\newdimen\strutht
\newdimen\strutdp
\strutht=\ht0
\strutdp=\dp0
\dimen0=\baselineskip
\advance\dimen0 by -\strutht
\advance\dimen0 by -\strutdp
\divide\dimen0 by 2
\advance\strutht by \dimen0
\advance\strutdp by \dimen0
\def\strut{\vrule width 0pt height \strutht depth \strutdp}%
%
\def\tab{%
  \kern 0.75ex
  \color{\r{ltgrey_fg}}{\vrule width 0.5pt}%
  \kern -0.5pt
  \kern -0.75ex
  \kern 3ex
}%
%
\def\kwd  #1{\color{\r{  blue_fg}}{#1}}% def, return, break, error
\def\flow #1{\color{\r{ amber_fg}}{#1}}% repeat, if, then, match
\def\liter#1{\color{\r{   red_fg}}{#1}}% numeric
\def\const#1{\color{\r{ lilac_fg}}{#1}}% empty, eof
\def\func #1{\color{\r{ green_fg}}{#1}}%
\def\var  #1{{\ifw #1}}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\centerline{\lcap8 UNIVERSAL CHESS INTERFACE}
\centerline{\lcap4 november 2024 • draft}

@@

\vskip-0.5\parskip The final position (\ot{\it P}\isub{N} if \s{moves} is
specified, otherwise {\it P}\sub{0}\ot) is the position \term{described} by
the \s{position} message. The collection {\it P}\isub{0}, {\it P}\isub{1},
\ellipsis, {\it P}\sub{{\subit N}−1} is the \term{history} of the \s{position}
message. If \s{moves} is not specified, the history is an empty collection.

\minor At any point in time, the history and position described by the
\s{position} message most recently dequeued by the engine are the \term{current
position history} and \term{current position}, respectively. If the engine has
not dequeued any \s{position} messages, the current position history and current
position are undefined. A \s{position} message is said to \term{set} the current
position history and current position.

@@

\minor A \term{search contextualizer} is a

\minor A \term{search limit} is a

@@

% + 999 999 999 999 999 999 fits within  0   to +2⁶⁴−1
% ± 999 999 999             fits within −2³¹ to +2³¹−1
% + 9999                    fits within  0   to +2¹⁶−1

% -?(0|[1−9][0−9]{0,8})   wtime btime winc binc
%      [1−9][0−9]{0,3}    mate movestogo (cannot be zero)
%    0|[1–9][0–9]{0,3}    depth
%    0|[1−9][0−9]{0,8}    movetime
%    0|[1−9][0−9]{0,17}   nodes

@@

\minor In some states, the engine may dequeue a message. If the engine dequeues
a message, the message must be well-formed and have an allowed message type
given the state of the engine.
\vskip-\parskip
\greybox{The word “may” is not meant to imply that standard input is ever
required to be empty, but meant to indicate that in some states the engine
is not allowed to read standard input (or at least must behave as if it has
not read standard input).}

@@

\minor The client must enqueue a \s{position} message at least once before it
enqueues a \s{go} message for the first time after enqueuing \s{uci}.
\vskip-\parskip
\bluebox{Engines are strongly recommended, if a \s{position} message is not
sent before the first \s{go} message, to behave as if \s{position startpos}
had been sent.}
\vskip-\parskip
\greybox{Note that the client is not required to send \s{isready} before
\s{go}.}
\vskip-\parskip
\greybox{Note that the client may send \s{ucinewgame} after \s{position}
before \s{go} without sending any additional intervening \s{position} message.}
\vskip-\parskip
\greybox{Note that \s{ucinewgame} does not reset the current board state nor
does it reset the move history. (Only \s{position} changes the current board
state and move history.)}

@@

\vskip\parskip
\begin
format("let @read_byte! : File_Descriptor -> Byte | EOF")
\end

\vskip\parskip
\begin
format(%q{
def withoutCR('msg : List(Byte)) : List(Byte)
  if empty?('msg) or last('msg) ≠ `0d then return 'msg
  return withoutLast('msg)
})
\end

\vskip\parskip
\begin
format(%q{
def read_message!(mut 'fd : File_Descriptor) : List(Byte) | EOF
  'msg <- empty
  repeat
    match read_byte!('fd)
      eof => return (\pt if empty?('msg) then eof else 'msg\pt)
      `0a => return withoutCR('msg)
      'val => append!('msg, 'val\pt)
})
\end

\vskip\parskip
\begin
format(%q{
def get_token!(mut 'msg : List(Byte)) : List(Byte) | None
  repeat
    if empty?('msg) then return none
    if first('msg) ≠ `20 then break
    removeFirst!('msg)
  'tok <- empty
  repeat
    append!('tok, first('msg))
    removeFirst!('msg)
    if empty?('msg) then return 'tok
    if first('msg) = `20 then break
  return 'tok
})
\end
